---
title: "mmstat4"
author: 
- name: "Sigbert Klinke" 
  email: sigbert@hu-berlin.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 4
vignette: > 
  %\VignetteIndexEntry{mmstat4} 
  %\VignetteEngine{knitr::rmarkdown} 
  \usepackage[utf8]{inputenc} 
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r , echo=FALSE}
suppressPackageStartupMessages(library(mmstat4))
```

# General

The aim of this package is to make teaching materials for different statistics courses available to students. This includes R, R Markdown, data and PDF files as well as Shiny Apps.

The central function `ghopen` can be used to

* Open Shiny Apps, R and R Markdown files in RStudio,
* load data files and
* display HTML and PDF files.

```{r}
library("mmstat4")
ghget("dummy")   # downloads a small repository
ghquery("bank2") # find data file 
x <- ghopen("data/BANK2.sav")
str(x)
```

If you choose to open a non-data set file then invisibly the full path to the file is returned.

## `ghget`, `ghset`

`ghset` set a key to a Zipfile of a GitHub repository. If it runs interactively then
you are asked if you want to install it locally to reduce the number of downloads. 
Otherwise it ia installed to a temporary directory, see `tempdir()`.
`ghget` downloads the GitHub repository if necessary. 

```{r, eval=FALSE}
library("mmstat4")
ghset("test", "https://github.com/sigbertklinke/mmstat4.dummy/archive/refs/heads/main.zip")
ghget("test") # if not yet installed
```

Note that you should call `ghget` or `ghset` whenever the `library` is called.

## `ghlist`, `ghquery`

With `ghlist` you can get a list of unique (short) names for all files or a subset based on a regular expression `pattern` in the repository

```{r}
str(ghlist())
ghlist("\\.pdf$") # get all PDF files
```

With `ghquery` you can query the list of unique (short) names for all files based on the overlap distance.

```{r}
ghlist("bnk")  # pattern = "bnk
ghquery("bnk") # nearest string matching to "bnk"
```


## Options

Several options can bet set with `options`

* `mmstat.dir` directory where the downloaded repository is stored, default `tempdir()`
* `mmstat.repo` link to GitHub directory, default `https://github.com/sigbertklinke/mmstat4/archive/refs/heads/main.zip`
* `mmstat.ext.prg` file extensions for a file to open in RStudio, default `c('', 'r', 'rmd', 'ma', 'py')`
* `mmstat.ext.doc` file extensions for a file to open ivia `browseURL`, default `c('html', 'pdf')`

# Skripte Dr. Sigbert Klinke, HU Berlin

## Basic statistics I+II (in german)

* Slides, exercises, solution and formulary for Statistics I+II (in german)

Mathematische Grundlagen - Einführung - Grundbegriffe - Univariate Verteilungen - Parameter  univariater Verteilungen - Bivariate Verteilungen - Parameter bivariater Verteilungen - Regressionanalyse - Zeitreihenanalyse - Indexzahlen - Wahrscheinlichkeitsrechnung - Zufallsvariablen - So lügt man mit Statistik - Wichtige Verteilungsmodelle - Stichprobentheorie - Statistische Schätzverfahren - Regressionsmodell - Konfidenzintervalle - Statistische Testverfahren - Parameterische Tests - Nichtparametrische Tests

```{r, eval=FALSE}
ghget("mmstat4")
ghopen("Statistik.pdf")
ghopen("Aufgaben.pdf")
ghopen("Loesungen.pdf")
ghopen("Formelsammlung.pdf")
```

* Slides for Data Analysis I+II

General - R - Basics and data generation - Test and estimation theory - Parameter of
distributions - Distribution - Transformations - Robust statistics - Missing values -
Subgroup analysis - Correlation and association - Multivariate graphics - Principal component analysis - Exploratory factor analysis - Reliability - Cluster analysis -
Regression analysis - Linear regression - Nonparametric regression - Classification and regression trees - Neural networks

```{r, eval=FALSE}
ghget("mmstat4")
ghopen("Statistik.pdf")
ghopen("Aufgaben.pdf")
ghopen("Loesungen.pdf")
ghopen("Formelsammlung.pdf")
```

## Data analysis I+II

# Skripte Prof. Bernd Rönz, HU Berlin (in german)

## Computergestützte Statistik I mit SPSS 10 (2001)

Einführung - Entdeckung und Identifikation von Ausreißern - Prüfung der Verteilungsform von Variablen - Parametervergleiche bei unbhängigen Stichproben - Anhänge A-D, Literaturverzeichnis, Index

```{r, eval=FALSE}
ghget("mmstat4")
ghopen("cs1.pdf")
```

## Computergestützte Statistik II mit SPSS 10 (2000)

Vorwort - Überprüfung von Zusammenhängen - Regressionsanalyse - Reliabilitäts- und Homogenitätsanalyse von Konstrukten - Anhänge A-H, Literaturverzeichnis, Stichwortverzeichnis

```{r, eval=FALSE}
ghget("mmstat4")
ghopen("cs2_roenz.pdf")
```

## Generalisierte lineare Modelle mit SPSS 10 (2001)

Einführung - Verallgemeinerte lineare Modelle (generalized linear models, GLM) - Modellierung binärer Daten - Das multinomiale Logit Modell - Modellierung multinomialer Daten (log-lineare Modelle) - Literaturverzeichnis, Index 

```{r, eval=FALSE}
ghget("mmstat4")
ghopen("glm.pdf")
```
