---
title: "mmstat4"
author: 
- name: "Sigbert Klinke" 
  email: sigbert@hu-berlin.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
vignette: > 
  %\VignetteIndexEntry{mmstat4} 
  %\VignetteEngine{knitr::rmarkdown} 
  \usepackage[utf8]{inputenc} 
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r , echo=FALSE}
suppressPackageStartupMessages(library("mmstat4"))
```

# General

The aim of this package is to provide teaching materials for various statistics courses for students.
The materials (R code snippets, Shiny apps, HTML and PDF documents, datasets, ...) are stored in a GitHub repository. GitHub allows downloading the repository as a ZIP file, see in the repository under the `Code` button (`Download ZIP`). `mmstat4` works with this ZIP file, but you can also use any of your own ZIP files.

In my courses I assume that all R programs run in a freshly started R, i.e. there are no path dependencies, all necessary libraries are loaded in the R program and so on. My repositories contain not only the example programs for the students, but also the programs I use to generate images and tables, and also the Shiny Apps I show.

# Using a ZIP file or repository

## `ghget`

A ZIP file or repository can be stored locally or in the internet.
A user defined key has to be given for the location of the ZIP file

```{r, eval=FALSE}
ghget(dummy="https://github.com/sigbertklinke/mmstat4.dummy/archive/refs/heads/main.zip")
```

Three repositories are predefined: `hu.data`, `hu.stat` and `dummy`. You can retrieve them via

```{r, eval=FALSE}
ghget('dummy')
ghget('hu.stat')
ghget('hu.data')
ghget()           # uses hu.data
```

`ghget` downloads the ZIP file, saves it to a temporary location and unpacks it. For non-temporary locations, see the FAQ.

## Full and short names for files

In addition, unique short names, related to the zip file content, are generated from the path components.

After unpacking the ZIP file, unique short names are generated for these files.

```{r}
ghget('dummy')
gd <- ghdecompose(ghlist(full.names=TRUE))
head(gd)
```

The file name is split into four parts. The last two parts, `minpath` and `filename`, are used to create short names:

1. the short name for `/tmp/RtmpXXXXXX/mmstat4.dummy-main/LICENSE` is `LICENSE`. There was no other file named `LICENSE` in the ZIP file. Therefore, it is sufficient to address this file in the ZIP file. 
2. the short name for `/tmp/RtmpXXXXXX/mmstat4.dummy-main/data/BANK2.sav` is `data/BANK2.sav`. There is another file called `BANK2.sav` in the ZIP file, but to address it uniquely, `data/BANK2.sav` is sufficient for this file in the ZIP file (the other is `dbscan/BANK2.sav`). Currently, no check is made whether two files with identical basenames are also identical in content.

```{r}
ghlist("BANK2", full.names=TRUE) # full names
ghlist("BANK2")                  # short names
```

## `ghopen`, `ghload`, `ghsource`

The short names (or the full names) can be used to work with the files

```{r, eval=4}
x <- ghload("data/BANK2.sav")          # load data via rio::import
ghopen("univariate/example_ecdf.R")    # open file in RStudio editor
ghsource("univariate/example_ecdf.R")  # execute file via source
ghlist("example_ecdf")                 # "univariate/" was unnecessary
```

## `ghlist`, `ghquery`

With `ghlist` you can get a list of unique (short) names for all files or a subset based on a regular expression `pattern` in the repository

```{r}
str(ghlist())     # get all short names
ghlist("\\.pdf$") # get all short names of PDF files
```

With `ghquery` you can query the list of unique (short) names for all files based on the overlap distance.

```{r}
ghlist("bnk")  # pattern = "bnk
ghquery("bnk") # nearest string matching to "bnk"
```

## `ghfile`, `ghpath`, `ghdecompose`

`ghfile` tries to find a unique match for a given file and returns the full path. If there is no unique match, an error is returned with some possible matches.

`ghdecompose` builds a data frame and decomposes the full names of the files into 

* `commonpath` the path part which is the same for all files,
* `uniquepath` the path part that is unique for all files,
* `minpath` the minimum path part, so that all files are uniquely addressable,
* `filename` the base name of the file, and
* `source` the input for shortpath.

The short names for the files are built from the components `minpath` and `filename`.

`ghpath` builds up the short name with various path components from a `ghdecompose` object.

```{r}
ghfile('data/BANK2.sav')
ghget(local=system.file("zip", "mmstat4.dummy.zip", package="mmstat4"))
fnf <- ghlist(full.names=TRUE)
dfn <- ghdecompose(fnf)
head(dfn)
head(ghpath(dfn))
```

## RStudio addins

The package comes with two RStudio addins (see under `Addins -> MMSTAT4`):

* _Open a file from a zip file_ (`ghopenAddin`), which gives access to the unzipped zip file and opens the selected file in an RStudio editor window.

* _Execute a Shiny app from a zip file_ (`ghappAddin`), which extracts all directories containing Shiny apps and opens the selected app in a web browser (using the default browser).

# Preparations for a ZIP file

Currently there are the following routines to support R code snippets:

* `Rlibs`, which extracts all `library` and `require` calls from the R code snippets and returns a frequency table of the packages called.

```{r}
ghget(local=system.file("zip", "mmstat4.dummy.zip", package="mmstat4"))
files <- ghlist(pattern="*.R$", full.names = TRUE)
head(Rlibs(files), 30)
```

* `Rsolo`, which checks that each R code snippet runs smoothly in a freshly started R. 

```{r, eval=FALSE}
# just check the last files from the list 
# Note that the R console will show more output (warnings etc.)
Rsolo(files, start=435)  
```

Note: there is also an error message if the necessary libraries are not installed!

# Frequently asked questions

## How can I force a reload of a zip file?

```{r, eval=FALSE}
ghget("dummy", .force=TRUE)
```

## How can I store a zip file permanently?

```{r, eval=FALSE}
ghget("dummy", .tempdir=FALSE)        # install non-temporarily
ghget("dummy", .tempdir="~/mmstat4")  # install non-temporarily to ~/mmstat4
ghget("dummy", .tempdir=TRUE)         # install again temporarily

```

Note: If a repository was installed permanently and you switch back to temporarily storage then the downloaded files will not be deleted.

## How can I find all directories with Shiny apps?

```{r, eval=FALSE}
ghget("dummy", .tempdir=TRUE)
ghlist(pattern="/(app|server)\\.R$")
ghopen("dbscan") # open the app
```

## How can I find all `csv` data files?

```{r}
ghget("dummy", .tempdir=TRUE)
ghlist(pattern="\\.csv$", ignore.case=TRUE, full.names=TRUE)
# use mmstat4::ghload for importing
ghlist(pattern="\\.csv$")
pechstein <- ghload("pechstein.csv")
str(pechstein)
```

# Default repositories

The package has three default repositories: `dummy`, `hu.stat`, and  `hu.data`. 

```{r, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| Repository       | Size          | ZIP file location |
| :--------------- | :-------------| :--------|
| `dummy`          | 3 MB          | `https://github.com/sigbertklinke/mmstat4.dummy/archive/refs/heads/main.zip`  |
| `hu.stat`        | 29 MB         | `https://github.com/sigbertklinke/mmstat4.data/archive/refs/heads/main.zip`   |
| `hu.data`        | 31 MB         | `https://github.com/sigbertklinke/mmstat4.stat/archive/refs/heads/main.zip`   |
"
cat(tabl) 
```

`dummy` is small subsample of `hu.stat` and `hu.data` which is intended for examples and test purposes. 

## Lecture Notes Sigbert Klinke, HU Berlin

### Basic statistics I+II (in german)

Mathematische Grundlagen - Einführung - Grundbegriffe - Univariate Verteilungen - Parameter  univariater Verteilungen - Bivariate Verteilungen - Parameter bivariater Verteilungen - Regressionanalyse - Zeitreihenanalyse - Indexzahlen - Wahrscheinlichkeitsrechnung - Zufallsvariablen - So lügt man mit Statistik - Wichtige Verteilungsmodelle - Stichprobentheorie - Statistische Schätzverfahren - Regressionsmodell - Konfidenzintervalle - Statistische Testverfahren - Parameterische Tests - Nichtparametrische Tests

```{r, eval=FALSE}
ghget("hu.stat")
ghopen("Statistik.pdf")
ghopen("Aufgaben.pdf")
ghopen("Loesungen.pdf")
ghopen("Formelsammlung.pdf")
```

### Data analysis

General - R - Basics and data generation - Test and estimation theory - Parameter of
distributions - Distribution - Transformations - Robust statistics - Missing values -
Subgroup analysis - Correlation and association - Multivariate graphics - Principal component analysis - Exploratory factor analysis - Reliability - Cluster analysis -
Regression analysis - Linear regression - Nonparametric regression - Classification and regression trees - Neural networks

```{r, eval=FALSE}
ghget("hu.data")
ghopen("dataanalysis.pdf")
```

## Lecture Notes Bernd Rönz, HU Berlin (in german)

### Computergestützte Statistik I mit SPSS 10 (2001)

Einführung - Entdeckung und Identifikation von Ausreißern - Prüfung der Verteilungsform von Variablen - Parametervergleiche bei unbhängigen Stichproben - Anhänge A-D, Literaturverzeichnis, Index

```{r, eval=FALSE}
ghget("hu.data")
ghopen("cs1_roenz.pdf")
```

### Computergestützte Statistik II mit SPSS 10 (2000)

Vorwort - Überprüfung von Zusammenhängen - Regressionsanalyse - Reliabilitäts- und Homogenitätsanalyse von Konstrukten - Anhänge A-H, Literaturverzeichnis, Stichwortverzeichnis

```{r, eval=FALSE}
ghget("hu.data")
ghopen("cs2_roenz.pdf")
```

### Generalisierte lineare Modelle mit SPSS 10 (2001)

Einführung - Verallgemeinerte lineare Modelle (generalized linear models, GLM) - Modellierung binärer Daten - Das multinomiale Logit Modell - Modellierung multinomialer Daten (log-lineare Modelle) - Literaturverzeichnis, Index 

```{r, eval=FALSE}
ghget("hu.data")
ghopen("glm_roenz.pdf")
```
